---
title: 'Final Project: Part 1'
author: "Cynthia Nickerson"
date: "June 19, 2019"
output: word_document
---

```{r}
library(tidyverse)
library(mice)
library(VIM)
library(ranger)
library(caret)
library(rpart)
library(rattle)
library(RColorBrewer)
library(rpart)
library(GGally)
library(ggcorrplot)
```

```{r}
rain <- read_csv("rain.csv")
View(rain)
```

```{r}
str(rain)
```

We are starting with 28,003 observations and 20 variables. There are several areas with missing data. See below.

```{r}
summary(rain)
```

```{r}
vim_plot = aggr(rain, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
```

I'm going to delete the rows with NA rather than try to impute the character records. 
```{r}
rainv2 <- rain %>%
  drop_na()
```

Now we have 13,887 observations with 20 variables.  

```{r}
vim_plot = aggr(rainv2, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)
```

Let's look at the relationship between the variables.  
```{r}
ggcorr(rainv2,lable = TRUE)
```

It is not unsurprising that there is a relationship between the 9am and 3pm temperatures and the max temperature.  IT's also unsurprising that there is a relatioship between windspeed, humidity, and pressure as the day passes. I'd like to focus on the relationship between Rainfall (since we are trying to predict rain) and Humidity.  

9am
```{r}
ggplot(rainv2,aes(x=Rainfall, y = Humidity9am)) + geom_point()
```

At 3pm
```{r}
ggplot(rainv2,aes(x=Rainfall, y = Humidity3pm)) + geom_point()
```

We can roughly see that the higher humidity, the better the chances are of rainfall.  At first glance this appears more so in the 9am data.

I'm going to convert some of these fields to factors before moving forward. 

```{r}
rainv3 <- rainv2 %>%
  mutate(RainToday = as.factor(RainToday))%>%
  mutate(RainToday =fct_recode(RainToday, "No"="0", "Yes"="1")) %>%
  mutate(RainTomorrow =as.factor(RainTomorrow))%>%
  mutate(RainTomorrow =fct_recode(RainTomorrow, "No"="0", "Yes"="1")) %>%
  mutate(MinTemp = as.factor(MinTemp)) %>%
  mutate(MaxTemp = as.factor(MaxTemp)) %>%
  mutate(Rainfall = as.factor(Rainfall)) %>%
  mutate(WindGustSpeed = as.factor(WindGustSpeed)) %>%
  mutate(WindSpeed9am = as.factor(WindSpeed9am)) %>%
  mutate(WindSpeed3pm = as.factor(WindSpeed3pm)) %>%
  mutate(Humidity9am = as.factor(Humidity9am)) %>%
  mutate(Humidity3pm = as.factor(Humidity3pm)) %>%
  mutate(Pressure9am = as.factor(Pressure9am)) %>%
  mutate(Pressure3pm = as.factor(Pressure3pm)) %>%
  mutate(Cloud9am = as.factor(Cloud9am)) %>%
  mutate(Cloud3pm = as.factor(Cloud3pm)) %>%
  mutate(Temp9am = as.factor(Temp9am)) %>%
  mutate(Temp3pm = as.factor(Temp3pm))


```

```{r}
ggplot(rainv3, aes(x=RainTomorrow, fill = RainToday))+ geom_bar()+theme_bw()
```

```{r}
t1 <- table(rainv3$RainTomorrow, rainv3$RainToday)
prop.table(t1, margin =2)
```

While RainToday is a predictor of RainTomorrow, I found the relationship with rainfall and humidity more curious. .  